---
title: "Homework #7, Part A"
author: "Jules Ford"
date: "2023-03-08"
output: html_document
---

**Problem 1**

I need to copy-paste the code from the homework and run them to get a bit familiar with them!

This is to read in the file:
```{r eval=FALSE, include=TRUE}
setwd("~/Documents/Comp Bio/FordBio381/docs")
z <- read.table("MyDataFile.csv",header=TRUE,sep=",")
str(z)
summary(z)
```

This is to open the libraries:
```{r}
library(ggplot2) # for graphics
library(MASS) # for maximum likelihood estimation
```

This is to read in the data vector:
```{r}
# quick and dirty, a truncated normal distribution to work on the solution set

z <- rnorm(n=3000,mean=0.2)
z <- data.frame(1:3000,z)
names(z) <- list("ID","myVar")
z <- z[z$myVar>0,]
str(z)
summary(z$myVar)
```

This is for plotting the data into a histogram:
```{r}
p1 <- ggplot(data=z, aes(x=myVar, y=..density..)) +
  geom_histogram(color="grey60",fill="cornsilk",size=0.2) 
print(p1)
```

This is creating an empirical density curve:
```{r}
p1 <-  p1 +  geom_density(linetype="dotted",size=0.75)
print(p1)
```

This generates maximum likelihood parameters for the normal curve:
```{r}
normPars <- fitdistr(z$myVar,"normal")
print(normPars)
str(normPars)
normPars$estimate["mean"] # note structure of getting a named attribute
```

This generates a curve for the normal probability density:
```{r}
meanML <- normPars$estimate["mean"]
sdML <- normPars$estimate["sd"]

xval <- seq(0,max(z$myVar),len=length(z$myVar))

 stat <- stat_function(aes(x = xval, y = ..y..), fun = dnorm, colour="red", n = length(z$myVar), args = list(mean = meanML, sd = sdML))
 p1 + stat
```


This adds a curve for exponential probability:
```{r}
expoPars <- fitdistr(z$myVar,"exponential")
rateML <- expoPars$estimate["rate"]

stat2 <- stat_function(aes(x = xval, y = ..y..), fun = dexp, colour="blue", n = length(z$myVar), args = list(rate=rateML))
 p1 + stat + stat2
```

This adds a curve for uniform probability density:
```{r}
stat3 <- stat_function(aes(x = xval, y = ..y..), fun = dunif, colour="darkgreen", n = length(z$myVar), args = list(min=min(z$myVar), max=max(z$myVar)))
 p1 + stat + stat2 + stat3
```

This adds a curve for gamma probability density:
```{r}
gammaPars <- fitdistr(z$myVar,"gamma")
shapeML <- gammaPars$estimate["shape"]
rateML <- gammaPars$estimate["rate"]

stat4 <- stat_function(aes(x = xval, y = ..y..), fun = dgamma, colour="brown", n = length(z$myVar), args = list(shape=shapeML, rate=rateML))
 p1 + stat + stat2 + stat3 + stat4
```

And this generates a new plot showing the beta probabiltiy density:
```{r}
pSpecial <- ggplot(data=z, aes(x=myVar/(max(myVar + 0.1)), y=..density..)) +
  geom_histogram(color="grey60",fill="cornsilk",size=0.2) + 
  xlim(c(0,1)) +
  geom_density(size=0.75,linetype="dotted")

betaPars <- fitdistr(x=z$myVar/max(z$myVar + 0.1),start=list(shape1=1,shape2=2),"beta")
shape1ML <- betaPars$estimate["shape1"]
shape2ML <- betaPars$estimate["shape2"]

statSpecial <- stat_function(aes(x = xval, y = ..y..), fun = dbeta, colour="orchid", n = length(z$myVar), args = list(shape1=shape1ML,shape2=shape2ML))
pSpecial + statSpecial
```

Alright! So that runs through everything once, running the entire sample code. Now, I'm going to run through all of this again, but with my own data set.
