---
title: "Homework5"
author: "Jules Ford"
date: "2023-02-15"
output: html_document
---

This is my Homework 5 for Computational Biology!

*Problem #1*


```{r}
myVec <- c(3:10)
n_dims <- sample(x=myVec, size=1)
print(n_dims)
selectedVal <- n_dims

# Create a vector of consecutive integers from 1 to n_dims^2
vec1 <- c(1:(n_dims)^2)
print(vec1)

# Use the sample function to randomly reshuffle these values.
vec2 <- sample(vec1)
print(vec2)

# create a square matrix with these elements.
matrix1 <- matrix(data=vec2, nrow=selectedVal, ncol=selectedVal)

# print out the matrix.
print(matrix1)

# find a function in r to transpose the matrix.
# the t() function does this!
tMat <- t(matrix1)

# print it out again and note how it has changed.
print(tMat)
# everything flipped on the diagonal!

# calculate the sum and the mean of the elements 
# in the first row and then the last row.

# sum of first row
sum(tMat[1,])

# mean of first row
mean(tMat[1,])

# sum of last row
sum(tMat[selectedVal,])

# mean of last row
mean(tMat[selectedVal,])

# read about the eigen()function and use it on your matrix
calc <- eigen(tMat, symmetric = TRUE, only.values=FALSE, EISPACK = FALSE)
print(calc)
# all of the elements of $values range widely
# all of the elements of $vectors are very low and range between -1 and 1

# dig in with the typeof() function to figure out their type.
typeof(calc)
# They are a list!

# if have set your code up properly, 
# you should be able to re-run it and create a matrix of 
# different size because n_dims will change

# I didn't get it the first time, but then I turned
# the value into variables and it worked!
```

For this problem, we had to randomly generate a number, square it, and then make a matrix using the squared value and modify the matrix. I had some trouble with it at first, but it was cool to try out some of the new functions! I used the `matrix()`
function to turn the sampled vector into a matrix. Then I used the `t()` function to transpose it, and took sums and averages of the first and last rows using functions like `sum(tMat[1,])` to access all of the values in the first row. By the end, I realized I didn't make it able to run with any generated number (I first generated 7 and plugged in 7 to the functions), so I made the variable `selectedVal` and substituted it where it needed to go.



*Problem #2*

```{r}
# Create a list with the following named elements:
# I will call it bigList

# my_matrix, which is a 4 x 4 matrix filled with random uniform values
my_matrix <- matrix(data = runif(n=16), nrow = 4, ncol = 4)

# my_logical which is a 100-element vector of TRUE or FALSE values. 
#Do this efficiently by setting up a vector of random values and then applying an inequality to it.
my_logical <- (runif(100) < 0.5)

# my_letters, which is a 26-element vector of all the lower-case letters in random order.
my_letters <- sample(letters)

bigList <- list(my_matrix, my_logical, my_letters)
print(bigList)

#Then, complete the following steps:

# create a new list, which has the element[2,2] from the matrix, 
# the second element of the logical vector, and the second element of the letters vector.
newList <- list(my_matrix[[2,2]], my_logical[2], my_letters[2])

# use the typeof() function to confirm the underlying data types of each component in this list
typeof(newList[[1]]) # this is a double
typeof(newList[[2]]) # this is a logical
typeof(newList[[3]]) # this is a character

# combine the underlying elements from the new list into a single atomic vector with the c() function.
# what is the data type of this vector?
finalVector <- c(typeof(newList[[1]]), typeof(newList[[2]]), typeof(newList[[3]]))
typeof(finalVector) # this is a character
```

For this problem we had to make a list of three different components using the `list()` function. Then, we had to access items from that list and put it into a new list, which required using `[]` and `[[]]` to confirm their types. Finally, we used `c()` to produce one last list composed of the types found from the prior list.

*Problem #3*

```{r}
# Create a data frame with the two variables (= columns) and 26 cases (= rows) below:

# call the first variable my_unis and fill it with 26 random uniform values from 0 to 10
my_unis <- runif(n = 26, min = 0, max = 10)

# call the second variable my_letters and fill it with 26 capital letters in random order.
my_letters <- sample(LETTERS)
  
# Creating the initial data frame:
firstFrame <- data.frame(my_unis, my_letters)
print(firstFrame)

# for the first variable, use a single line of code in R to select 4 random rows 
# and replace the numerical values in those rows with NA.
firstFrame[sample(1:26, size = 4),1] <- NA
print(firstFrame)

#for the first variable, write a single line of R code to identify which rows have the missing values.
firstFrame[!complete.cases(firstFrame),]


#re-order the entire data frame to arrange the second variable in alphabetical order

firstFrame$my_letters # This line is for my sake; it says the original order of the letters

order(firstFrame$my_letters) # This line is for my sake; it says the original row numbers of the letters of the alphabet (before firstFrame was alphabetized, "A" was in row 24)

firstFrame[order(firstFrame$my_letters), ] # This properly reorders things!

#calculate the column mean for the first variable.
workableFrame <- firstFrame[complete.cases(firstFrame),] # making a new data frame with only the rows that have values
print(mean(workableFrame[,1])) # calculating and printing the mean
```

This problem asks for us to produce a data frame that has two variables. I made the first variable, `my_unis`, with the `runif()` function and set the minimums and maximums. For the second variable I used the `sample()` function with `LETTERS` as the argument, which is known in R to provide the letters of the alphabet, capitalized, in a randomized order. I used these variables and the `data.frame()` function to make the data frame. Then, I randomly assigned 4 rows in it the value `NA`, and used the `!complete.cases()` function to name the rows where `NA` values can be found. I used `order()` to sort the data frame alphabetically from the `my_letters` variable. Finally, I used `complete.cases()` once more to grab the rows from `firstFrame` that don't have values, assigned it to `workableFrame`, and calculated its mean.
