---
title: "Homework #7, Part B"
author: "Jules Ford"
date: "2023-03-08"
output: html_document
---

**Problem 2**

I am using the dataset from this research article: "A rodent herbivore reduces its predation risk through ecosystem engineering" from the Dryad website. [Click here to go to the web page on Dryad that describes this research.](https://datadryad.org/stash/dataset/doi:10.5061%2Fdryad.cvdncjt57)

```{r}
voleData <- read.table("hwk7Data.csv",header=TRUE,sep=",")
str(voleData)
summary(voleData)
```

```{r}
library(ggplot2) # for graphics
library(MASS) # for maximum likelihood estimation
```

This is a histogram for the initial plant volume:
```{r}
p1 <- ggplot(data=voleData, aes(x=Initial.plant.volume..m3., y=..density..)) +
  geom_histogram(color="grey60",fill="cornsilk",size=0.2) 
print(p1)
```
This is the empirical density curve:
```{r}
p1 <-  p1 +  geom_density(linetype="dotted",size=0.75)
print(p1)
```


This gets the maximum likelihood parameters for the normal distribution:
```{r}
normPars <- fitdistr(voleData$Initial.plant.volume..m3.,"normal")
print(normPars)
str(normPars)
normPars$estimate["mean"] # note structure of getting a named attribute
```


This plots the normal probability density:
```{r}
meanML <- normPars$estimate["mean"]
sdML <- normPars$estimate["sd"]

xval <- seq(0,max(voleData$Initial.plant.volume..m3.),len=length(voleData$Initial.plant.volume..m3.))

 stat <- stat_function(aes(x = xval, y = ..y..), fun = dnorm, colour="red", n = length(voleData$Initial.plant.volume..m3.), args = list(mean = meanML, sd = sdML))
 p1 + stat
```

This plots the exponential probability density:

```{r}
expoPars <- fitdistr(voleData$Initial.plant.volume..m3.,"exponential")
rateML <- expoPars$estimate["rate"]

stat2 <- stat_function(aes(x = xval, y = ..y..), fun = dexp, colour="blue", n = length(voleData$Initial.plant.volume..m3.), args = list(rate=rateML))
 p1 + stat + stat2
```
This plots the uniform probability density:
```{r}
stat3 <- stat_function(aes(x = xval, y = ..y..), fun = dunif, colour="darkgreen", n = length(voleData$Initial.plant.volume..m3.), args = list(min=min(voleData$Initial.plant.volume..m3.), max=max(voleData$Initial.plant.volume..m3.)))
 p1 + stat + stat2 + stat3
```

This plots the gamma probability distribution:
```{r}
gammaPars <- fitdistr(voleData$Initial.plant.volume..m3.,"gamma")
shapeML <- gammaPars$estimate["shape"]
rateML <- gammaPars$estimate["rate"]

stat4 <- stat_function(aes(x = xval, y = ..y..), fun = dgamma, colour="brown", n = length(voleData$Initial.plant.volume..m3.), args = list(shape=shapeML, rate=rateML))
 p1 + stat + stat2 + stat3 + stat4
```

And last but not least, this plots the beta probability density:
```{r}
pSpecial <- ggplot(data=voleData, aes(x=Initial.plant.volume..m3./(max(Initial.plant.volume..m3. + 0.1)), y=..density..)) +
  geom_histogram(color="grey60",fill="cornsilk",size=0.2) + 
  xlim(c(0,1)) +
  geom_density(size=0.75,linetype="dotted")

betaPars <- fitdistr(x=voleData$Initial.plant.volume..m3./max(voleData$Initial.plant.volume..m3. + 0.1),start=list(shape1=1,shape2=2),"beta")
shape1ML <- betaPars$estimate["shape1"]
shape2ML <- betaPars$estimate["shape2"]

statSpecial <- stat_function(aes(x = xval, y = ..y..), fun = dbeta, colour="orchid", n = length(voleData$Initial.plant.volume..m3.), args = list(shape1=shape1ML,shape2=shape2ML))
pSpecial + statSpecial
```

**Problem 3**

Honestly, I think that the uniform probability density is the one that best fits the data. This was surprising!



