---
title: "Homework #10"
author: "Jules Ford"
date: "2023-04-19"
output: html_document
---
**Problem 1**
```{r}
vec <- c(8, 4, 0, 10, 900, 4, 0)

####################################
# FUNCTION: zeroCalculator
# packages: none
# purpose: determine the number of zeros inside a numeric vector
# input: a numeric vector
# output: the number of zeros inside the numeric vector
#----------------------------
crazyZeroCalculator <- function(vec) {
  counter <- 0
  betterVec <- paste(vec, collapse="")
  betterVec <- strsplit(betterVec, split="")
  for (i in seq_along(betterVec[[1]])) {
    if (betterVec[[1]][i] == "0") {
     counter <- counter + 1
    }
  }
  return(counter)
}

check <- crazyZeroCalculator(vec) 
print(check)
```
This problem requires us to take a vector and determine the number of zeroes in that vector. To make this function, I used the `paste()` function so I could collapse the vector into a single character string with no spaces. Then, I used the `strsplit()` function to separate each character from that string into its own character string object. Then, using a for loop, I checked whether that object was the character string `"0"` and if it was, I added 1 to the counter variable (the running total of zeroes in that vector).
 
**Problem 2**
```{r}
newVec <- c(8, 4, 0, 10, 900, 4, 0, 0, 20)

####################################
# FUNCTION: shortZeroCalc
# packages: none
# purpose: use subsetting to determine whether each value in a vector contains at least one 0
# input: a numeric vector
# output: the number of items inside the numeric vector that contain at least one 0
#----------------------------
shortZeroCalc <- function(vector) {
  return(length(vector[vector %% 10 == 0]))
}

numOfZeroes <- shortZeroCalc(newVec)
numOfZeroes
```
This problem tasks us with shortening our first answer into by using subsetting rather than looping to accomplish the same task. In the first problem I went the trickier route of building a function that could count every single zero (so it could detect that 900 had two 0s). In simplifying this version, because I didn't go the same route of turning everything into character strings, I settled for the counter being able to detect whether a given item had at least one 0. 
 
**Problem 3**
```{r}
####################################
# FUNCTION: matrixMaker
# packages: none
# purpose: generate a matrix given its row and column
# input: 2 integers, reprsenting rows and columns in a matrix
# output: a matrix with the inputted dimensions, where each element is the product of the given row and column number
#----------------------------
matrixMaker <- function(rows, cols) {
  mat <- matrix(data=NA, nrow=rows, ncol=cols) # this makes the matrix
  for (i in 1:nrow(mat)) {
    for (j in 1:ncol(mat)) { # I'm going to loop through each row and column
      mat[i,j] <- (i * j) # and assign the product of that point in the loop (i represents the row # and j represents the column #)
    }
  }
  return(mat)
}

test <- matrixMaker(8,4)
```

This problem tasks us with making a function that generates a matrix based on an input amount of rows and columns, where we have to fill each part of the matrix with the product of the row and column. To accomplish this, I used a for loop to iterate through each row and column in the matrix. For each part of the matrix, I multiplied the row and column number (denoted by i and j within the loop) to get the proper product!

**Problem 4**
**4a**
**4b**
**4c**
**4d**