---
title: "Homework #4"
author: "Jules Ford"
date: "2023-02-08"
output: html_document
---
This is Homework #4 for Computational Biology.



**Problem #1**

The code:

```{r}
x <- 1.1
a <- 2.2
b <- 3.3

# Problem 1a
z <- x^a^b
print(z)

# Problem 1b
z <- (x^a)^b
print(z)

# Problem 1c
z <- (3*x^3 + 2*x^2 + 1)
print(z)
```


This problem asks for us to assign values to x, a, and b, and then to assign expressions to z using those variables and computing them. To solve the problem, I used the assignment operator `<-` to assign different values to variables and the `print()` function to display the calculated values.



**Problem #2**

The code: 

```{r}
# Problem 2a
c(1:8, 7:1)

# Problem 2b

# I started out with this code:
# c(rep(1,1), rep(2,2), rep(3,3), rep(4,4), rep(5,5))
# which did the job but was too clunky
# Here's a prettier version:

rep(1:5, c(1, 2, 3, 4, 5))

# Problem 2c

# Similarly, I started out with this code:
# c(rep(5,1), rep(4,2), rep(3,3), rep(2,4), rep(1,5))
# which also did the job but was too clunky
# Here's a prettier version:

rep(5:1, c(1, 2, 3, 4, 5))

```

This problem asks for us to come up with code to print different sequences officially. For 2a, I used a `seq()` shortcut to count to 8 then back to 1. For problems 2b and 2c, I used the `rep()` function to specify the range of numbers in the sequence, and a `c()` function to specify the amount of times each value in the sequence should be repeated.



**Problem #3**

The code:

```{r}
set.seed(0)
# Produce x and y
x <- runif(n=1)
y <- runif(n=1)
# Print x and y
c(x, "is x", y, "is y")
# Produce r and theta
r <- sqrt(x^2 + y^2)
theta <- atan(y/x)
# Print r and theta
c(r, "is r", theta, "is theta")
```


For this problem I had to generate random numbers for an (x, y) Cartesian coordinate and I needed to convert those coordinates to polar coordinates. I started by setting the seed, then randomly generating values for x and y with the `runif()` function. After, I printed x and y so I could check my work. Using x and y, I set their conversion equations to r and theta, and printed the results!



**Problem #4**

The Code:

```{r}
# Start of the line
queueInitial <- c("sheep", "fox", "owl", "ant")
print(queueInitial)
# Serpent joins line in the back
queueA <- append(queueInitial, "serpent")
print(queueA)
# Sheep gets on the ark and leaves
queueB <- queueA[ !queueA == "sheep"]
print(queueB)
# Donkey enters line straight at the front
queueC <- append(queueB, "donkey", after = 0)
print(queueC)
# Serpent gets impatient and leaves
queueD <- queueC[ !queueC == "serpent"]
print(queueD)
# Owl gets bored and leaves
queueE <- queueD[ !queueD == "owl"]
print(queueE)
# Aphid enters line and cuts ant
queueF <- append(queueE, "aphid", after = 2)
print(queueF)
queueF[3]
```

For this problem, we needed to update a line as different animals entered and left. We used `append()` to add different animals to the line and `queueX[ !queueX == "y"]` to get rid of items in the vector.



**Problem #5**

The code:

```{r}
bigVec <- c(1:100) # I start by making a vector called bigVec with numbers 1-100
print(bigVec) # printing it out to confirm that it works

endVec <- bigVec[(bigVec %% 2 != 0) & (bigVec %% 3 != 0) & (bigVec %% 7 != 0)] # I make a new vector that's assigned to the selection of values from bigVec

print(endVec) # I print out the vector at the end
```

This problem asks for us to create a vector for numbers 1-100 that aren't divisible by 2, 3, or 7 using arithmetic operators. `&` means and, `|` means or, and `5 %% 4` gives 1 as a remainder. To solve this problem, I started by making one vector with the `c()` function that includes numbers from 1 to 100. Then, I made a smaller vector that would only include the desired values from the first vector. To select the numbers I want, I used `bigVec[]` to select values from that vector. Within the `[]`, I said `(bigVec %% 2 != 0)`, which takes each number from `bigVec` and divides it by 2, and checks its remainder. I want the values that have remainders not equal to 0, because if its remainder is 0 then it's divisible by 2. I repeated this process with `(bigVec %% 3 != 0)` and `(bigVec %% 7 != 0)`, separating each with the `&` operator which tells the program that I don't want any values divisible by 3 or 7 either. I finished it off by printing `endVec` to check my work!